\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

This thesis is the result of the first few years of my academic journey, but I
have been afforded the want and the resilience to get to this point by a great
number of people. Here, I would like to acknowledge a non-exhaustive list of
those people.

First and most importantly, I offer my utmost gratitude to my supervisors, Dr
Jonathan Gillard and Dr Vincent Knight. Your support, encouragement and guidance
over the course of this project have been invaluable. You have made a great team
for me, and your distinct brand of tutelage has instilled in me equal measures
of curiosity, determination and scepticism. I would also like to thank
Mr Kendal Smith and those at Cwm Taf Morgannwg for giving me the opportunity to
work on a project that has inspired great motivation and intrigue for me.

I would like to thank my family, who have been a source of inspiration and
support for me throughout my studies. Thank you to my parents, Susan and John,
who have always pushed me to exceed myself. Thank you to my siblings for
facilitating a healthy amount of competition growing up. Finally, I offer a
special thanks to my older brother, Matthew, for looking out for me when I have
needed it most.

To my dearest friends, Noah Atkin, Jessica Lockwood, Cameron Pearson, and Edward
Priest, I offer my sincere thanks and admiration. Your unwavering kindness,
warmth and solace has been vital to keeping my head above water at the worst of
times, and has filled me with a great sense of joy otherwise.

Finally, I turn my attention to my academic home for the last six or so years,
Cardiff University. I wish to thank my peers in the School of Mathematics,
especially Emma Aspland, Lorenzo De Biase, Nikoleta Glynatsi, Emily O'Riordan,
Geraint Palmer, Chris Seaman and \'{A}lvaro Torras Casas, for your stimulating
conversation, resolute solidarity and welcome distraction. Lastly, thank you to
the staff of the School for maintaining a nourishing environment in which to
study and work.
