\chapter*{Dissemination}
\addcontentsline{toc}{chapter}{Dissemination}

\section*{Publications}

\begin{itemize}
    \item H. Wilde, V. Knight, and J. Gillard. \textbf{Evolutionary dataset
        optimisation: learning algorithm quality through evolution}.
        \emph{Applied Intelligence}, 50(4):1172–1191,
        2020.~\doi{10.1007/s10489-019-01592-4}.
    \item H. Wilde, V. Knight, and J. Gillard. \textbf{Matching: A Python
        library for solving matching games}. \emph{Journal of Open Source
    Software}, 5(48):2169, 2020.~\doi{10.21105/joss.02169}.
\end{itemize}

\section*{In preparation}

\begin{itemize}
    \item H. Wilde. \textbf{A review of current literature covering the
        intersections of clustering, algorithm evaluation and healthcare
        modelling}. Available on GitHub at
        \github{daffidwilde/literature-review}.
    \item H. Wilde, V. Knight, and J. Gillard. \textbf{A novel initialisation
        based on hospital-resident assignment for the \(k\)-modes algorithm}.
        Preprint available at \arxiv{2002.02701} --- \emph{submitted to
        Knowledge and Information Systems}.
    \item H. Wilde, V. Knight, J. Gillard, and K. Smith. \textbf{Segmentation
        analysis and the recovery of queuing parameters via the Wasserstein
        distance: a study of administrative data for patients with chronic
        obstructive pulmonary disease}. Preprint available at \arxiv{2008.04295}
        --- \emph{under review at the European Journal of Operational Research}.
    \item H. Wilde, V. Knight, and L. Roach. \textbf{Automatic final-year
        project allocation in a School of Biosciences}.
\end{itemize}

\section*{Talks and posters}

Unless otherwise stated, the following items are available at
\github{daffidwilde/talks}.

\begin{itemize}
    \item \textbf{Benchmark datasets are not the only option when understanding
        algorithm performance (poster)}. \emph{SIAM UKIE Annual Meeting},
        University of Edinburgh. 2020. Available at
        \github{daffidwilde/siam-ukie}. 
    \item \textbf{Learning algorithm quality through evolution}. \emph{Welsh
        Mathematics Colloquium}, Gregynog Hall (Powys). 2019.
    \item \textbf{Parallelisation in Python}. \emph{Advanced Python Workshop},
        Cardiff University. 2019.
    \item \textbf{Evolutionary dataset optimisation: data synthesis and
        algorithm quality}. \emph{Data Science Campus Seminar Series}, Office
        for National Statistics. 2019.
    \item \textbf{Understanding variation with clustering}. \emph{NHS Wales
        Modelling Collaborative}, Cardiff. 2019.
    \item \textbf{What is Dask?} \emph{PyDiff}, Cardiff. 2017.
\end{itemize}

\section*{Software development}

\begin{itemize}
    \item \edo: a Python library for generating artificial datasets through
        evolution. The main application of the library is to better understand
        an algorithm's quality by studying the generated datasets. Available
        at \github{daffidwilde/edo}. \textbf{Contributions:} main developer.
    \item \edolab: an open source framework for conducting reproducible
        experiments with \edo. Available at \github{daffidwilde/edolab}.
        \textbf{Contributions:} main developer.
    \item \matching: a Python library for facilitating and solving various
        matching games. Available at \github{daffidwilde/matching}.
        \emph{Contributions:} main developer.
    \item \kmodes: a Python implementation of the \(k\)-modes and
        \(k\)-prototypes algorithms. Available at \github{nicodv/kmodes}.
        \textbf{Contributions:} implemented epoch cost recording and matching-based
        initialisation.
    \item \ciw: a simulation library for open queuing networks. Available at
        \github{CiwPython/Ciw}. \emph{Contributions: helped refactor
        distributions into Python objects, added to documentation}.
\end{itemize}
